Informing forest carbon inventories under the Paris Agreement using the Global Forest Carbon Database (ForC v4.0)

### 2024-01-29 doc/manuscript/ForC-EFDB manuscript.Rmd
minor

# Introduction

@@

The goal of this publication is to facilitate submission of forest C estimates from scientific studies to EFDB.
We document the process of submitting data to EFDB from the Global Forest Carbon Database, ForC ([https://forc-db.github.io/](https://forc-db.github.io/)), which is the largest collection of published estimates of forest C stocks, increments, and annual fluxes [@anderson-teixeira_forc_2018; @anderson-teixeira_carbon_2021; @anderson-teixeira_forcdb_2023], including data ingested from individual publications and relevant databases, including the Global Reforestation Opportunity Assessment (GROA) database [@cook-patton_mapping_2020, database doi: 10.5281/zenodo.3983644], and the global soil respiration database [SRDB-V5, @bond-lamberty_global_2010; @jian_restructured_2021].
We (1) map common scientific forest C estimation methods and definitions to those used by the IPCC;
(2) develop a semi-automated process for preparing ForC data for submission to EFDB;
and (3) assess the data in ForC relevant to EFDB and records that have been submitted to date;
We conclude with recommendations as to how the scientific community can better provide useful data for forest C inventories under the Paris Agreement.

# Materials and Methods

Major steps for submission of data from ForC to EFDB included
(1) mapping ForC into EFDB, including aligning ForC and IPCC definitions (summarized in `r appendix_IPCC_definitions`),
(2) revising ForC v3.0 to support semi-automated submissions to EFDB, yielding ForC v4.0 (detailed in `r appendix_ForC_updates`), and
(3) submitting data to EFDB. 
These steps are detailed below.

## Mapping ForC to EFDB

With input from the IPCC's Technical Support Unit and referencing IPCC guidance [@ipcc_good_2003; @ipcc_2006_2006; @ipcc_2019_2019], we mapped ForC fields into EFDB fields, as summarized in `r table_ForCfieldmapping`.
For the majority of fields, contents of the field in ForC was copied directly into an EFDB field, either as the only contents of that field or as part of a composite record. <!--For example, ten ForC fields describing site location, climate, and edaphic properties all mapped into the EFDB field *Region/Regional conditions* (`r table_ForCfieldmapping`).-->
For some fields, simple conditional logic was used to populate EFDB fields based on ForC records.<!--For example, for stock variables presented in the original publication in units of dry organic matter mass (as opposed to C), several greenhouse gasses (CO~2~, CO, CH~4~, NO, NO~2~, N~2~O) were entered in the EFDB field indicating the greenhouse gases to which the record could be pertinent (*Gases* field) because these values could be used in calculations of greenhouse gas emissions from biomass burning [@ipcc_2006_2006]; otherwise, the only pertinent greenhouse gas would be CO~2~.--> 
In cases where original studies did not present 95% confidence intervals (required by IPCC when available) but did present standard error or n and standard deviation, we calculated the 95% confidence intervals.<!--and populated the EFDB field with this information (noting the calculation in the EFDB field *Comments from Data Provider*)-->
There were two cases in which more complex mapping was required: (1) mapping of C cycle variables (detailed below in **Carbon cycle variables**) and (2) land classification (detailed below in **Land classification**).

### Carbon cycle variables

@@

### Land classification

@@

## Updating ForC

@@

## Submission of ForC data to EFDB

@@

# Results

## ForC v4.0 contents

@@

## Data submissions to EFDB

@@

# Discussion

@@

## Database needs

@@

## Data collection and analysis needs

@@

## Data reporting needs

@@

# Conclusions

@@

# Acknowledgements
  We gratefully acknowledge the substantial contributions of Valentyna Slivinska and Sandro Federici for collaboration on the conception, design, and technical review of this project.
  Thank you to all researchers who collected and published the data contained in ForC, and to all research assistants and collaborators who have contributed to building the database.
  Thank you to Avni Malhotra for helpful comments on an earlier draft of this manuscript.
  Funding for this study was provided by the Smithsonian (Forest Global Earth Observatory, Smithsonian Working Land and Seascapes); a Bezos Earth Fund grant to the Nature Conservancy, with a sub-grant to NZCBI; and the Institute for Global Environmental Strategies.

# Author Contribution
  KAT and VH conceived and designed the project; VH wrote the scripts for database management, data submission to EFDB, and the analyses presented here; MW, TR, and RBM added and reviewed ForC data, BBL and SCP contributed large databases to ForC (EFDB and GROA, respectively); CP provided methodological expertise; KAT, VH, and MW prepared the first draft of the manuscript; all authors reviewed the results and approved the final version of the manuscript.

# Data Availability Statement 
All code and data are openly available. The ForC database and associated code are available via the ForC repository within the ForC-db organization on GitHub (https://github.com/forc-db/ForC), and the version used here (ForC v4.0) is archived in Zenodo [@anderson-teixeira_forcdb_2023, DOI: 10.5281/zenodo.8020861]. The data and code associated with data submission to EFDB and preparation of this manuscript are available via the the IPCC-EFDB-integration repository within the ForC-db organization on GitHub  (https://github.com/forc-db/IPCC-EFDB-integration) and archived in Zenodo (DOI: 10.5281/zenodo.8021474).

\clearpage
# References

>>>>

# Introduction {-}

@@

The goal of this publication is to facilitate submission of forest C estimates from scientific studies to EFDB.
We document the process of submitting data to EFDB from the Global Forest Carbon Database, ForC ([https://forc-db.github.io/](https://forc-db.github.io/)), which is the largest collection of published estimates of forest C stocks, increments, and annual fluxes [@anderson-teixeira_forc_2018; @anderson-teixeira_carbon_2021; @anderson-teixeira_forcdb_2023], including data ingested from individual publications and relevant databases, including the Global Reforestation Opportunity Assessment (GROA) database [@cook-patton_mapping_2020, database doi: 10.5281/zenodo.3983644], and the Global Soil Respiration Database [SRDB-V5, @bond-lamberty_global_2010; @jian_restructured_2021].
We (1) map common scientific forest C estimation methods and definitions to those used by the IPCC;
(2) develop a semi-automated process for preparing ForC data for submission to EFDB;
and (3) assess the data in ForC relevant to EFDB and records that have been submitted to date.
We conclude with recommendations as to how the scientific community can better provide useful data for forest C inventories under the Paris Agreement.

# Materials and Methods {-}

Major steps for submission of data from ForC to EFDB included
(1) mapping ForC into EFDB, including aligning ForC terms and concepts with those defined by IPCC guidelines (summarized in `r appendix_IPCC_definitions`),
(2) revising ForC v3.0 to support semi-automated submissions to EFDB, yielding ForC v4.0 (detailed in `r appendix_ForC_updates`), and
(3) submitting data to EFDB. 
These steps are detailed below.

## 1. Mapping ForC to EFDB {-}

With input from the IPCC's Technical Support Unit and referencing IPCC guidance [@ipcc_good_2003; @ipcc_2006_2006; @ipcc_2019_2019], we mapped ForC fields into EFDB fields, as summarized in `r table_ForCfieldmapping`.
For the majority of fields, contents of the field in ForC mapped directly into an EFDB field, either as the only contents of that field or as part of a composite record. <!--For example, ten ForC fields describing site location, climate, and edaphic properties all mapped into the EFDB field *Region/Regional conditions* (`r table_ForCfieldmapping`).-->
For some fields, simple conditional logic was used to populate EFDB fields based on ForC records.<!--For example, for stock variables presented in the original publication in units of dry organic matter mass (as opposed to C), several greenhouse gasses (CO~2~, CO, CH~4~, NO, NO~2~, N~2~O) were entered in the EFDB field indicating the greenhouse gases to which the record could be pertinent (*Gases* field) because these values could be used in calculations of greenhouse gas emissions from biomass burning [@ipcc_2006_2006]; otherwise, the only pertinent greenhouse gas would be CO~2~.--> 
For example, in cases where original studies did not present 95% confidence intervals (required by IPCC when available) but did present standard error or n and standard deviation, we calculated the 95% confidence intervals.<!--and populated the EFDB field with this information (noting the calculation in the EFDB field *Comments from Data Provider*)-->
There were two cases in which more complex mapping was required: (1) mapping of C cycle variables (detailed below in **Carbon cycle variables**) and (2) land classification (detailed below in **Land classification**).

### Carbon cycle variables {-}

@@

### Land classification {-}

@@

## 2. Updating ForC {-}

@@

## 3. Submission of ForC data to EFDB {-}

@@

# Results {-}

## ForC v4.0 contents {-}

@@

## Data submissions to EFDB {-}

@@

# Discussion {-}

@@

## Database needs {-}

@@

## Data collection and analysis needs {-}

@@

## Data reporting needs {-}

@@

# Conclusions {-}

@@

# Acknowledgements {-}
  We gratefully acknowledge the substantial contributions of Valentyna Slivinska and Sandro Federici for collaboration on the conception, design, and technical review of this project.
  Thank you to all researchers who collected and published the data contained in ForC, and to all research assistants and collaborators who have contributed to building the database.
  Thank you to Avni Malhotra for helpful comments on an earlier draft of this manuscript.
  Funding for this study was provided by the Smithsonian (Forest Global Earth Observatory, Smithsonian Working Land and Seascapes); a Bezos Earth Fund grant to the Nature Conservancy, with a sub-grant to NZCBI; and the Institute for Global Environmental Strategies.

# Author Contribution {-}
  KAT and VH conceived and designed the project; VH wrote the scripts for database management, data submission to EFDB, and the analyses presented here; MW, TR, and RBM added and reviewed ForC data, BBL and SCP contributed large databases to ForC (EFDB and GROA, respectively); CP provided methodological expertise; KAT, VH, and MW prepared the first draft of the manuscript; all authors reviewed the results and approved the final version of the manuscript.

# Data Availability Statement  {-}
All code and data are openly available. The ForC database and associated code are available via the ForC repository within the ForC-db organization on GitHub (https://github.com/forc-db/ForC), and the version used here (ForC v4.0) is archived in Zenodo [@anderson-teixeira_forcdb_2023, DOI: 10.5281/zenodo.8020861]. The data and code associated with data submission to EFDB and preparation of this manuscript are available via the the IPCC-EFDB-integration repository within the ForC-db organization on GitHub  (https://github.com/forc-db/IPCC-EFDB-integration) and archived in Zenodo (DOI: 10.5281/zenodo.8021474).

\clearpage
# References {-}
